Manuscript Heat Map Script

### 2025-09-03 ABP_heatmap_for_Cece_20241018.Rmd
Removing relative file path and adding missing pacakge to p_load
Added YSI data to heatmaps

pacman::p_load(tidyverse, akima, reshape2, gridExtra, colorRamps, RColorBrewer, ggpubr, geomtextpath, scales)

@@

 EDI_CTD <-
  read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/14/0432a298a90b2b662f26c46071f66b8a")%>%
   filter(DateTime>ymd_hms("2018-01-01 00:00:00") & DateTime<ymd_hms("2023-12-31 00:00:00"))%>%
   filter(Reservoir=="BVR"|Reservoir=="FCR")%>%
   filter(Site==50)%>%
   select(-starts_with("Flag"))%>%
  mutate(Year=year(DateTime))%>%
  filter(Depth_m>0)

@@

CTD_casts <- EDI_CTD |>
  select(Reservoir, DateTime, Year)|>
  unique()

@@

@@

      if(nrow(bvr)>0){
        interp <- interp2xyz(interp(x = as.numeric(bvr$Date), y= bvr$Depth_m, z= unlist(bvr[[var]]),
                                        xo = seq(min(bvr$Date), max(bvr$Date), "day"),
                                        yo = seq(min(bvr$Depth_m), max(bvr$Depth_m), 
                                                 by = 0.1), 
                                        duplicate = "mean"),data.frame = T)
        interp2<-interp%>%
    mutate(Date=as.Date(x, origin = "1970-01-01"),
           Reservoir=Res)

@@

```{r plotting}

@@

make_heat_map <- function(
    data,
    sample_dates,
    turnovers, 
    y_lim,
    custom_heatmap_palette,
    y_units,
    output_file){

@@

# change the plots dimensions  
jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "black") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="black", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank())+
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette)+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing.x = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)
  # labs(color = "")+
  # geom_point(data = sensor_depth_plot, 
  #            aes(x = Date, y = Sensor_depth), 
  #            shape = 23, fill = "black", color = "white", size  =2.5)+
  # geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  # scale_linetype_manual(values = c("solid","84"),
  #                       limits = c("Thermocline","Outtake pipe"),  
  #                       name= NULL)+

@@

wer <- data.frame(
  data= c(rep("EDI_CTD",12)),
  var = c(rep("Temp_C",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

@@

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts,
  turnovers = turnover, 
  y_lim = c(-1, 30),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "Water temperature (°C)",
  output_file = "figures/SedTraps_CTD_waterTemp_2018_2023.jpeg"
)

@@

wer <- data.frame(
  data= c(rep("EDI_CTD",12)),
  var = c(rep("DO_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

@@

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts,
  turnovers = turnover, 
  y_lim = c(-1,16),
  custom_heatmap_palette = rev(blue2green2red(60)),
  y_units = "Dissolved Oxygen mg/L",
  output_file = "figures/SedTraps_CTD_DO_2018_2023.jpeg"
)

>>>>

pacman::p_load(tidyverse, akima, reshape2, gridExtra, colorRamps, RColorBrewer, ggpubr, geomtextpath, scales, rLakeAnalyzer)

@@

 EDI_CTD <-
  read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/14/0432a298a90b2b662f26c46071f66b8a")%>%
   filter(DateTime>ymd_hms("2018-01-01 00:00:00") & DateTime<ymd_hms("2023-12-31 00:00:00"))%>%
   filter(Reservoir=="BVR"|Reservoir=="FCR")%>%
   filter(Site==50)%>%
   select(-starts_with("Flag"))%>%
  mutate(Year=year(DateTime),
         Date = date(DateTime))%>%
  filter(Depth_m>0)

@@

CTD_casts <- EDI_CTD |>
  select(Reservoir, DateTime, Year)|>
  mutate(Date = date(DateTime)) %>% 
  distinct(Reservoir, Date, .keep_all = TRUE)

@@

# Make a vector of sampling days so CTD and handhelds do not overlap
  CTD_days <- EDI_CTD %>% 
    select(Date) %>% 
    unique()

# add in handhelds to supplement data gaps
  Handhelds  <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/198/13/e50a50d062ee73f4d85e4f20b360ce4f") %>% 
    filter(between(DateTime, as.Date('2018-01-01 00:00:00'),
                   as.Date('2023-12-31 23:59:59')),
           Reservoir == 'FCR' | Reservoir == 'BVR',
           Site == 50,
           Depth_m >= 0) %>% 
    arrange(DateTime)

'%!in%' <- function(x,y)!('%in%'(x,y))

Handhelds1 <- Handhelds %>% 
  mutate(Date = date(DateTime),
         Year = year(DateTime)) %>% 
  filter(Date %!in% CTD_days$Date) %>% 
  select(Reservoir, Site, DateTime, Depth_m, Temp_C, DO_mgL, Date, Year) %>% 
  #fixing leap day issue, shifting sample date by one day
  mutate(Date = if_else(Date == as.Date('2020-02-29'), 
                          as.Date('2020-02-28'), Date))

Handheld_days <- Handhelds1 |>
  select(Reservoir, Date, DateTime, Year)|>
  distinct(Reservoir, Date, .keep_all = TRUE)

# GEt the month and date for easier graphing. This is not the correct date.   
  Handheld_days$MonthDay <- format(Handheld_days$Date,"%m-%d")
  Handheld_days$MDay <- as.Date(Handheld_days$MonthDay, "%m-%d")

EDI_CTD_Handhelds <- EDI_CTD %>% 
  select(Reservoir, Site, DateTime, Depth_m, Temp_C, DO_mgL, Date, Year) %>% 
  rbind(Handhelds1) %>% 
  na.omit() %>% 
  arrange(Date)

@@

      if(nrow(bvr)>0){
        interp <- interp2xyz(interp(x = as.numeric(bvr$Date), y= bvr$Depth_m, z= unlist(bvr[[var]]),
                                        xo = seq(min(bvr$Date), max(bvr$Date), "day"),
                                        yo = seq(min(bvr$Depth_m), max(bvr$Depth_m), 
                                                 by = 0.5), 
                                        duplicate = "mean"),data.frame = T)
        interp2<-interp%>%
    mutate(Date=as.Date(x, origin = "1970-01-01"),
           Reservoir=Res)

@@

```{r heat map}

@@

make_heat_map <- function(
    data,
    sample_dates1,
    sample_dates2,
    turnovers,
    y_lim,
    custom_heatmap_palette,
    y_units,
    output_file){

@@

# change the plots dimensions  
jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates1, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "black") + #to mark cast dates 
  geom_point(data = sample_dates2, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "black") +
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="black", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank())+
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette)+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing.x = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)
  # labs(color = "")+
  # geom_point(data = sensor_depth_plot, 
  #            aes(x = Date, y = Sensor_depth), 
  #            shape = 23, fill = "black", color = "white", size  =2.5)+
  # geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  # scale_linetype_manual(values = c("solid","84"),
  #                       limits = c("Thermocline","Outtake pipe"),  
  #                       name= NULL)+

@@

wer <- data.frame(
  data= c(rep("EDI_CTD_Handhelds",12)),
  var = c(rep("Temp_C",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

@@

make_heat_map(
  data= temp_df,
  sample_dates1 = CTD_casts,
  sample_dates2 = Handheld_days,
  turnovers = turnover, 
  y_lim = c(-1, 30),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "Water temperature (°C)",
  output_file = "figures/SedTraps_waterTemp_2018_2023.jpeg"
)

@@

wer <- data.frame(
  data= c(rep("EDI_CTD_Handhelds",12)),
  var = c(rep("DO_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

@@

make_heat_map(
  data= temp_df,
  sample_dates1 = CTD_casts,
  sample_dates2 = Handheld_days,
  turnovers = turnover, 
  y_lim = c(-1,16),
  custom_heatmap_palette = rev(blue2green2red(60)),
  y_units = "Dissolved Oxygen mg/L",
  output_file = "figures/SedTraps_DO_2018_2023.jpeg"
)
